---
title: "Simple Weather App"
author: "Juan Tellez"
output: html_document
runtime: shiny
---


I was tasked with creating a simple app in Shiny that would provide weather information for a wide variety of cities. Weather data is collected via a web API from the website *Forecast.io*. The workhorse function here is `get_Weather` [reproduced below], which takes an API key, longitude and latitude for a selected city, and returns weather information that is later parsed and organized for visualization. 



```{r, echo = TRUE, eval= FALSE}
  get_Weather <- function(apiKey, lat, long)
{
  url = sprintf("https://api.forecast.io/forecast/%s/%s,%s", 
                apiKey, lat, long) # cite: http://bit.ly/1NPplsf
  
  web = GET(url = url) # pull from website
  web = content(web, as = "parsed")
  
  daily = web$daily # pull out daily forecasts
  
  output = lapply(daily$data, as.data.frame) %>% 
    ldply(. , data.frame) #turn each list into data frame and then row-bind them
  
```



In terms of the App itself, the user selects one of the cities on the list as well as a secondary weather characteristic they would like information about. The city is then geocoded using the `ggmap`
package and longitude and latitude are extracted. These are then fed to the `get_Weather` function alongside the API key. The output is then used to generate the plots. 

The app includes two conditional panels to generate either one plot or side-by-side plots according to whether the user prefers one kind of plot or another. 


```{r, echo=FALSE, message=FALSE}
  library(ggmap)
  library(httr)
  library(dplyr)
  library(shiny)
  library(gridExtra)
  library(plyr)
  library(data.table)
  
  # define cities
  cities <- c("Bogota", 
              "New York City",
              'New Delhi',
              'London',
              'Paris', 
              'Shanghai', 
              'Lagos', 
              'Istanbul',
              'Tokyo',
              'Moscow',
              'Rio de Janeiro',
              'Seoul',
              'Cairo',
              'Lima',
              'Tehran',
              'Baghdad', 
              'Dhaka',
              'Santiago',
              'Alexandria',
              'Cape Town',
              'Berlin', 
              'Kabul',
              'Madrid',
              'Buenos Aires',
              'Chicago', 
              'Miami', 
              'Dublin', 
              'Belfast')
  
  cities = sort(cities) # sort in alpha order
  cities <- as.character(cities) # make sure all character vector
  
  apiKey = "b60f5443971c2de6116a5558d3b1f842" # define API key
  
  
  # alternative weather options
  
  weather = c("Precipitation Intensity", 
              "Max Precipitation Intensity",
              "Probability of Precipitation", 
              "Dew Point", "Humidity","Wind Speed",
              "Wind Bearing", "Cloud Cover", "Pressure")
  
  # main panel of app
  
  mainPanel(
    
    
    br(),
    
    selectInput("input_city", label = "City to Forecast:",
                choices = cities, selected = 'NAIROBI'),
    
    checkboxInput("alternative", label = "More Weather Metrics", value = FALSE),
    
    conditionalPanel(
      condition = "input.alternative == true", 
      selectInput("alt_temp", label = "Alternative Weather Options:",
                   choices = weather, selected = weather[1])
    )
    
  
  )
  
  # default panel (low and highs at each day)
  conditionalPanel(
    condition = "input.alternative == false", 
    renderPlot({
    # Set a theme for gg
    theme_set(theme_minimal())
    
    # get Weather from Forecast.io
    get_Weather <- function(apiKey, lat, long)
  {
    url = sprintf("https://api.forecast.io/forecast/%s/%s,%s", 
                  apiKey, lat, long) # cite: http://bit.ly/1NPplsf
    
    web = GET(url = url) # pull from website
    web = content(web, as = "parsed")
    
    daily = web$daily # pull out daily forecasts
    
    output = lapply(daily$data, as.data.frame) %>% 
      ldply(., data.frame) #turn each list into data frame and then row-bind them
    
  #   # turn into datetime
    output$time <- as.POSIXct(output$time, origin = "1970-01-01", tz = "UTC")
    
    output$temperatureMinTime <- as.POSIXct(output$temperatureMinTime,
                                                origin = "1970-01-01", tz = "UTC")
    output$temperatureMaxTime <- as.POSIXct(output$temperatureMaxTime,
                                                origin = "1970-01-01", tz = "UTC")
    output$apparentTemperatureMinTime <- as.POSIXct(output$apparentTemperatureMinTime,
                                        origin = "1970-01-01", tz = "UTC")
    output$apparentTemperatureMaxTime <- as.POSIXct(output$apparentTemperatureMaxTime,
                                        origin = "1970-01-01", tz = "UTC")
    
    
    
    # Fix Names
  weathervars = c("precipIntensity", 
              "precipIntensityMax", 
              "precipProbability", 
              "dewPoint", 
              "humidity", 
              "windSpeed", 
              "windBearing",
              "cloudCover", 
              "pressure")
  
  newlabels =  c("Precipitation Intensity", 
              "Max Precipitation Intensity",
              "Probability of Precipitation", 
              "Dew Point", "Humidity","Wind Speed",
              "Wind Bearing", "Cloud Cover", "Pressure")
  
  
  output = data.table::setnames(output, old = weathervars, new = newlabels)
    
    
     return(output)
    }
    
    # pull long and lat for specified city
    locations = geocode(input$input_city, source = 'google')
    
    # get weather
    today = get_Weather(apiKey = apiKey, 
                lon = locations$lon, 
                lat = locations$lat)
    
    
    # make the plot
    
    tmp = ggplot()
    tmp = tmp + theme_minimal()
    tmp = tmp + geom_line(data = today,
                        aes(x=temperatureMinTime, 
                            y=temperatureMin,
                        color = "yellow"), size = 3) # stock color assignment
    tmp = tmp + geom_line(data = today, 
                        aes(x = temperatureMaxTime, 
                            y = temperatureMax,
                            color = "red"), size = 3) # stock color assignment
    
    tmp = tmp + labs(title = paste("Forecasted Temperature for", input$input_city, sep = " "), x = "", y = "")
    
    tmp = tmp + theme(legend.position = 'bottom')
    tmp = tmp + scale_color_manual(name="Temperature Forecast:", 
                                 values=c("#2171b5", "#bdd7e7"), 
                                 labels = c("High Temp.","Low Temp."))
    tmp = tmp + theme(plot.title = element_text(face = 'bold'))
    
    print(tmp)
  
  })
  )
    conditionalPanel(
      condition = "input.alternative == true",
      renderPlot({
        
        # Set a theme for gg
    theme_set(theme_bw())
    
    # get Weather from Forecast.io
    get_Weather <- function(apiKey, lat, long)
  {
    url = sprintf("https://api.forecast.io/forecast/%s/%s,%s", 
                  apiKey, lat, long) # cite: http://bit.ly/1NPplsf
    
    web = GET(url = url) # pull from website
    web = content(web, as = "parsed")
    
    daily = web$daily # pull out daily forecasts
    
    output = lapply(daily$data, as.data.frame) %>% 
      ldply(. , data.frame) #turn each list into data frame and then row-bind them
    
  #   # turn into datetime
    output$time <- as.POSIXct(output$time, origin = "1970-01-01", tz = "UTC")
    
    output$temperatureMinTime <- as.POSIXct(output$temperatureMinTime,
                                                origin = "1970-01-01", tz = "UTC")
    output$temperatureMaxTime <- as.POSIXct(output$temperatureMaxTime,
                                                origin = "1970-01-01", tz = "UTC")
    output$apparentTemperatureMinTime <- as.POSIXct(output$apparentTemperatureMinTime,
                                        origin = "1970-01-01", tz = "UTC")
    output$apparentTemperatureMaxTime <- as.POSIXct(output$apparentTemperatureMaxTime,
                                        origin = "1970-01-01", tz = "UTC")
    
    
    
     # Fix Names
  weathervars = c("precipIntensity", 
              "precipIntensityMax", 
              "precipProbability", 
              "dewPoint", 
              "humidity", 
              "windSpeed", 
              "windBearing",
              "cloudCover", 
              "pressure")
  
  newlabels =  c("Precipitation Intensity", 
              "Max Precipitation Intensity",
              "Probability of Precipitation", 
              "Dew Point", "Humidity","Wind Speed",
              "Wind Bearing", "Cloud Cover", "Pressure")
  
  
  output = data.table::setnames(output, old = weathervars, new = newlabels)
    
    
     return(output)
    }
    
    # pull long and lat for specified city
    locations = geocode(input$input_city, source = 'google')
    
    # get weather
    today = get_Weather(apiKey = apiKey, 
                lon = locations$lon, 
                lat = locations$lat)
    
    
    # make the plot
    
    tmp = ggplot()
    tmp = tmp + theme_minimal()
    tmp = tmp + geom_line(data = today,
                        aes(x=temperatureMinTime, 
                            y=temperatureMin,
                        color = "yellow"), size = 3)
    tmp = tmp + geom_line(data = today, 
                        aes(x = temperatureMaxTime, 
                            y = temperatureMax,
                            color = "red"), size = 3)
    tmp = tmp + labs(title = paste("Forecasted Temperature for", input$input_city, sep = " "), x = "", y = "")
    tmp = tmp + theme(legend.position = 'bottom')
    tmp = tmp + scale_color_manual(name="Temperature Forecast:", 
                                 values=c("#2171b5", "#bdd7e7"), 
                                 labels = c("High Temp.", "Low Temp."))
    tmp = tmp + theme(plot.title = element_text(face = 'bold'))
        
    tmp2 = ggplot()
    tmp2 = tmp2 + theme_minimal()
    tmp2 = tmp2 + geom_line(data = today, 
                            aes_string(x = "time", 
                                y = paste0("`", input$alt_temp, "`")), 
                            color = "#6baed6", size = 3)
    tmp2 = tmp2 + labs(title = paste("Forecasted", input$alt_temp, "in", 
                                     input$input_city, sep = " "),
                       x = "", y = "")
    tmp2 = tmp2 + theme(plot.title = element_text(face = 'bold'))
    grid.arrange(tmp, tmp2, ncol = 2)
      })
    )
  

```

